---
title: "253 project"
author: "Oriana Galasso and Ty Bruckner"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) #for plotting and summarizing
library(GGally) #for nice scatterplot matrix 
library(ggridges) #for joy/ridge plots
library(corrplot) #for basic correlation matrix plot
library(naniar) #for exploring missing values
library(pdp) #for partial dependence plots, MARS models
library(rpart.plot) #for plotting decision trees
library(vip) #for importance plots
library(janitor)
#making things look nice
library(lubridate) #for nice dates
library(knitr) #for nice tables
library(scales) #for nice labels on graphs
library(gridExtra) #for arranging plots
library(broom) #for nice model output
library(janitor) #for nice names

#data
library(ISLR) #for data
library(moderndive) #for data

#modeling
library(rsample) #for splitting data
library(recipes) #for keeping track of transformations
library(caret) #for modeling
library(leaps) #for variable selection
library(glmnet) #for LASSO
library(earth) #for MARS models
library(rpart) #for decision trees
library(randomForest) #for bagging and random forests

```

```{r}
cbb <- read_csv("cbb.csv")

#Split the data into training and test groups
set.seed(253) #for reproducibility
cbb_split <- initial_split(cbb, 
                             prop = .7)
cbb_train <- training(cbb_split)
cbb_test <- testing(cbb_split)
```

```{r}
head(cbb)
```

free throws attempted/field goals attempted 

## Tentative research question: Which stats or other predictors, such as conference,  are most important to predict NCAA tournament seed? 

```{r}
cbb%>%
  ggplot(aes(x=W,y=SEED))+
  geom_point()+
  geom_smooth()
```

Wins seem to coincide with a lower seed (better seed in the tournament). This makes sense.

```{r, fig.height=3}
cbb%>%
  ggplot(aes(x=W,y=SEED))+
  geom_point() +
  facet_wrap(~CONF) +
  geom_smooth(method = "lm")
```
Here we can see wins and seeds faceted by conference. We see that many conferences almost exclusively receive low seeds (non- power five conferences). 

```{r}
cbb%>%
  ggplot(aes(x=W))+
  geom_histogram() +
  facet_wrap(~CONF)
```


```{r}
cbb%>%
  ggplot((aes(x=TOR,y=W)))+
  geom_jitter() +
  geom_smooth(method = "lm")
```

This makes sense-- teams with a lot of turnovers that they commit have higher seeds (worse seeds).

```{r}
cbb%>%
  ggplot(aes(x=TORD,y=W))+
  geom_point()+
  geom_smooth()
```

Defensive turnover rate (turnovers you force the other team do make) does not seem like an effective predictor by itself.

```{r}
cbb%>%
  ggplot(aes(x=ADJ_T,y=W))+
  geom_jitter()+
  geom_smooth(method = "lm")
```

Adjusted Tempo does not seem to affect seed by itself. 

```{r}
cbb %>%
  ggplot(aes(x=`3P_O`, y=W)) +
  geom_jitter() +geom_smooth(method = "lm")
```

```{r}
cbb %>%
  ggplot(aes(x=`2P_O`, y=W)) +
  geom_jitter() +geom_smooth(method = "lm")
```

```{r}
cbb %>%
  ggplot(aes(x=FTR, y=W)) +
  geom_jitter() +geom_smooth(method = "lm")
```

```{r}
cbb %>%
  ggplot(aes(x=FTRD, y=W)) +
  geom_jitter() +geom_smooth(method = "lm")
```

```{r}
cbb %>%
  ggplot(aes(x=BARTHAG, y=W)) +
  geom_jitter() +geom_smooth()
```

Preds we're interested in: 

3PO
2PO
FTRD
TOR
BARTHAG

# Beginning modeling 

## Linear Model



```{r}
set.seed(253)
cbb_linear1 <- train(
  W ~ `2P_O` + `3P_O` + FTRD + TOR + BARTHAG,
  data = cbb_train, 
  method = "lm",
  trControl = trainControl(method = "cv", 
                           number = 5))
  na.action = na.omit


cbb_linear1$results #average RMSE

```

```{r}
head(cbb)
```

```{r}
set.seed(253)
cbb_linear2 <- train(
  W ~ ADJOE + ADJDE + BARTHAG + `EFG_O` + `EFG_D` + TOR + TORD + ORB + DRB + FTR + FTRD + `2P_O` + `2P_D` + `3P_O` +`3P_D`+ `ADJ_T`,
  data = cbb_train, 
  method = "lm",
  trControl = trainControl(method = "cv", 
                           number = 5))
  na.action = na.omit


cbb_linear2$results #average RMSE
```


```{r}
lambda_grid <- 10^seq(-3, 1, length = 100)

set.seed(253)
cbb_lasso <- train(
  W ~ .,
  data = cbb_train %>% select(-SEED, -POSTSEASON,-G,-TEAM,-CONF,-YEAR),
  method = "glmnet",
  trControl = trainControl(method="cv",number=5),
  tuneGrid = data.frame(alpha = 1, lambda = lambda_grid),
  na.action = na.omit
)
```

```{r}
cbb_lasso$results
```
```{r}
cbb_lasso$bestTune
```
```{r}
coefficients(cbb_lasso$finalModel,.003351603)
```
I am confused on our interpretation here.

We need to transform variables I think?
```{r}

set.seed(253)
knn_4 <- train(
  W ~ .,
  data = cbb_train%>%select(-SEED, -POSTSEASON,-G,-TEAM,-CONF,-YEAR),
  method = "knn",
  tuneGrid = data.frame(k = c(1,2,3,4,5,6)),
  trControl = trainControl(method = "cv", number = 5),
  na.action = na.omit
)
```

```{r}
knn_4$results
```


```{r}
cbb_train%>%
  mutate(pred_W = predict(knn_4, 
                        newdata = cbb_train))%>% ggplot(aes(x = (W), y = (pred_W))) +
  geom_point(size = .5, alpha = .5) +
  geom_abline(color = "purple") +
  geom_smooth(se = FALSE, color = "orange", size = .5) 
```



```{r}
set.seed(253)
cbb_tree <- train(
  W ~ ADJOE + ADJDE + `EFG_O` + `EFG_D` + TOR + TORD + ORB + DRB + FTR + FTRD ,
  data = cbb_train, 
  method = "rpart",
  trControl = trainControl(method = "cv",
                           number = 5),
    tuneGrid = data.frame(cp = 10^seq(-4, -2 , length = 20))
)

```

```{r}
cbb_tree$results
```
```{r}
cbb_tree$bestTune
```

```{r, fig.width=10, fig.height=10}
rpart.plot(cbb_tree$finalModel)
```

```{r}
vip(cbb_tree$finalModel, num_features = 12, bar = FALSE)
```


```{r}
set.seed(253)

cbb_ADJOE <- train(
  W ~ ADJOE,
  data = cbb_train, 
  method = "lm",
  trControl = trainControl(method = "cv", 
                           number = 5))
  na.action = na.omit


cbb_ADJOE$results #average RMSE

```


