---
title: "253 project"
author: "Oriana Galasso"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
```

```{r}
cbb <- read_csv("cbb.csv")
```

```{r}
cbb
```


Which stats are most important to predict tournament seed? 
```{r}
cbb%>%
  ggplot(aes(x=W,y=SEED))+
  geom_point()+
  geom_smooth()
```
Wins seem to coincide with a lower seed.
```{r}
cbb%>%
  ggplot(aes(x=CONF,y=SEED))+
  geom_point()
```
Some conferences are relegated to only recieving lower seeds.

```{r}
cbb%>%
  ggplot((aes(x=TOR,y=SEED)))+
  geom_point()+
  geom_smooth()
```
```{r}
cbb%>%
  ggplot(aes(x=TORD,y=SEED))+
  geom_point()+
  geom_smooth()
```
Defensive turnover rate does not seem like an effective predictor by itself.

```{r}
cbb%>%
  ggplot(aes(x=ADJ_T,y=SEED))+
  geom_point()+
  geom_smooth()
```

Adjusted Tempo does not seem to affect seed by itself. 


I tried to start the plots. We can add more too. We might need to rearrange the Postseason variable to make it make sense. 

