---
title: "253 project"
author: "Oriana Galasso and Ty Bruckner"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
```

```{r}
cbb <- read_csv("cbb.csv")
```

```{r}
head(cbb)
```


## Tentative research question: Which stats or other predictors, such as conference,  are most important to predict NCAA tournament seed? 

```{r}
cbb%>%
  ggplot(aes(x=W,y=SEED))+
  geom_point()+
  geom_smooth()
```

Wins seem to coincide with a lower seed (better seed in the tournament). This makes sense.

```{r, fig.height=3}
cbb%>%
  ggplot(aes(x=W,y=SEED))+
  geom_point() +
  facet_wrap(~CONF) +
  geom_smooth(method = "lm")
```
Here we can see wins and seeds faceted by conference. We see that many conferences almost exclusively receive low seeds (non- power five conferences). 

```{r}
cbb%>%
  ggplot((aes(x=TOR,y=SEED)))+
  geom_point() +
  geom_smooth(method = "lm")
```

This makes sense-- teams with a lot of turnovers that they commit have higher seeds (worse seeds).

```{r}
cbb%>%
  ggplot(aes(x=TORD,y=SEED))+
  geom_point()+
  geom_smooth()
```

Defensive turnover rate (turnovers you force the other team do make) does not seem like an effective predictor by itself.

```{r}
cbb%>%
  ggplot(aes(x=ADJ_T,y=SEED))+
  geom_point()+
  geom_smooth()
```

Adjusted Tempo does not seem to affect seed by itself. 

```{r}
cbb %>%
  ggplot(aes(x=`3P_O`, y=W)) +
  geom_point() +geom_smooth(method = "lm")
```

This is kind of confusing. It seems like a better 3 point percentage means teams have higher seeds (worse seeds). We should investigate further. 


